Research log

### 2024-01-07 docs/paper.Rmd
typo

abstract: |
  What we do depends on what we know. But sometimes we try to decouple our behavior from our knowledge so as to appear not to know what we really do know. Such pretense behavior requires understanding how we would behave with different knowledge, and so provides a window into counterfactual self-simulation. However, little research has characterized and evaluated pretense relative to non-pretense behavior. Here, in two large-scale, pre-registered experiments, we tested people’s ability to conceal knowledge in the context of information-seeking games (Battleship and Hangman). Participants played normal games (in which they began with no knowledge of the ships’ locations or the word’s identity), as well as ‘pretend’ games in which they were shown the game's full solution but were instructed to play *as though* they were ignorant. We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' paticipants failed to distinguish real players from pretenders. Nonetheless, pretense was characterized by pervasive differences, undetected by judges but revealed by follow-up analyses: pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules. Our findings thus reveal a powerful but imperfect ability to simulate decisions and behavior based on a counterfactual belief state, and open up a new approach to studying self-simulation.

@@

The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University. `E2.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation. They were selected based on their acceptance rate (\>95%) and for being native English speakers. The entire experiment took approximately 20 minutes to complete. Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r``E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).

>>>>

abstract: |
  What we do depends on what we know. But sometimes we try to decouple our behavior from our knowledge so as to appear not to know what we really do know. Such pretense behavior requires understanding how we would behave with different knowledge, and so provides a window into counterfactual self-simulation. However, little research has characterized and evaluated pretense relative to non-pretense behavior. Here, in two large-scale, pre-registered experiments, we tested people’s ability to conceal knowledge in the context of information-seeking games (Battleship and Hangman). Participants played normal games (in which they began with no knowledge of the ships’ locations or the word’s identity), as well as ‘pretend’ games in which they were shown the game's full solution but were instructed to play *as though* they were ignorant. We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' participants failed to distinguish real players from pretenders. Nonetheless, pretense was characterized by pervasive differences, undetected by judges but revealed by follow-up analyses: pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules. Our findings thus reveal a powerful but imperfect ability to simulate decisions and behavior based on a counterfactual belief state, and open up a new approach to studying self-simulation.

@@

The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University. `E2.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation. They were selected based on their acceptance rate (\>95%) and for being native English speakers. The entire experiment took approximately 20 minutes to complete. Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (``` r``E2.bonus$bonus%>%min() ``` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
